Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-10 popl24/intro.tex
checkpoint

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten, koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in blockchains \cite{ishai2009zero, lu2019honeybadgermpc, gao2022symmeproof, tomaz2020preserving}. MPC methods
have been developed by the cryptography community for many years,
while receiving the attention of the programming languages community
only relatively recently. In contrast, information flow security has
received significant attention in PL theory and practice especially
since the turn of the century \cite{1159651}, including a menagerie of
variants, enforcement mechanisms, and programming frameworks. Much of
this has been enabled by the unified metatheory of
\emph{hyperproperties} \cite{10.5555/1891823.1891830} that establishes
a common conceptual framework for reasoning about and implementing
systems with information flow security.

Our goal is to explore and establish connections between the security
model of MPC-- the \emph{real/ideal} aka \emph{simulator security}
model -- and trace-based hyperproperties, and to leverage these
connections to obtain automated enforcement mechanisms for MPC
protocol development. Currently, proof methods for MPC protocol
development are well-studied \cite{Lindell2017} but manual and 
thus tedious and error-prone, and are also non-standardized and
unfamiliar to most PL theorists. Therefore our exploration will
make both theoretical and practical contributions by bridging a gap between
information flow and simulator security methodologies.

\subsection{The Analysis Challenge of MPC}
\jnote{On its surface, security for MPC protocols seems similar to probabilistic noninterference. However, MPC security differs from noninterference in a subtle but fundamental way, and approaches for verifying noninterference cannot naturally extend to MPC security.}
MPC protocols involve communication
between a group of distributed participants called a \emph{federation}
that collaboratively compute and publish the result of some known
\emph{ideal functionality} $\idealf$, while keeping each party's input
``secret'', without the use of a trusted third party. This last part
is critical. For example, if we take $\idealf$ to be the majority vote
function, a protocol for computing $\idealf$ is MPC-secure if, given
any set of input votes, it correctly computes and publishes the voting
result but reveals no other information to the public or to other
participants. However, by publishing the result, some information
about individual votes may be implicitly declassified.  For example,
in the case of a majority vote in a federation of size 3, if the
motion carries and party 1 has voted no, then party 1 knows exactly
the votes of parties 2 and 3. This cannot be avoided due to the nature
of $\idealf$.

@@

In Section \ref{section-hyperprop-passive} we formalize real/ideal
security, and we define and discuss simple examples of MPC protocols
in Section \ref{section-minicat-examples} and more complex ones in
Section \ref{section-composition}. But an immediate and main point in
relation to security hyperproperties is that, due to the potential for
information release in MPC, simulator security is \emph{not} a strict
probabilistic noninterference or trace obliviousness property, as we
show in Section \ref{section-hyperprop-ni}- rather, the public output
allows and sets an upper bound on declassification.

@@

In Section \ref{section-minicat} we develop a new probabilistic programming
language $\minifed$ for defining synchronous distributed protocols
over the binary field. The syntax and semantics provides a succinct
account of \emph{views} and (synchronous) messaging between protocol
\emph{clients}. In this Section, we consider small example protocols
including oblivious transfer, communications encryption with one-time
pads, and additive secret sharing. In Section \ref{section-pmf} we
formulate \emph{program distributions}, a formalism for expressing
dependencies between secrets and views, useful for expressing
intensional characteristics of protocols and precisely quantifying
allowable information leakage in terms of protocol outputs. On this
basis we then formulate a novel hyperproperty of program execution
traces, called \emph{noninterference modulo output ($\NIMO$)} that we
demonstrate implies passive simulator security. In Section
\ref{section-bruteforce} we define a brute force mechanism for
verifying passive security in $\minifed$ protocols, that is also
amenable to HPC optimizations via translation of $\minifed$ protocols
into Datalog.

In Section \ref{section-metalang} we define a new metaprogramming
language $\metaprot$ that dynamically generates $\minifed$
protocols. It includes control and data structures and is able to
express logical protocol components, and enjoys a type safety result
that guarantees that generated protocols are semantically
well-defined. Our key analysis results are presented in Section
\ref{section-composition}, where we show how intensional,
automatically certifiable properties of program components can be
composed to obtain whole-program MPC security. We consider the GMW and
YGC protocols as examplars, both of which use \emph{circuit}
abstractions comprising \emph{gates} in arbitrarily large
configurations, and show how gates can be automatically certified to
ensure security in circuits that use them in a manner that scales to
any circuit size. This approach is extensible, in that new gate
implementations can be automatically certified to guarantee safe
inclusion in circuit libraries.

In summary, our analysis takes a ``just do it'' approach- rather than
using, e.g., traditional security type or linear type disciplines to
prevent ``bad things'' from happening, we base our analysis on precise
quantification of intensional protocol behavior and especially
dependencies between secret inputs, public outputs, and adversarial
views of protocols. We ameliorate scalability issues associated with
fine-grained analysis by manually establishing compositional
properties inspired by probabilistic separation logic
\cite{barthe2019probabilistic}, obtaining a middle-ground, scalable
``proof assistant style'' approach enabled by MPC circuit paradigms.

>>>>

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. MPC methods have been
developed by the cryptography community for many years, while
receiving the attention of the programming languages community more
recently. In contrast, information flow security has received
significant attention in PL theory and practice especially since the
turn of the century \cite{1159651}, including a menagerie of variants,
enforcement mechanisms, and programming frameworks. Much of this has
been enabled by the unified metatheory of \emph{hyperproperties}
\cite{10.5555/1891823.1891830}, including noninterference, that
establishes a common conceptual framework for reasoning about and
implementing systems with information flow security.

On its surface, security for MPC protocols seems similar to
probabilistic noninterference. However, MPC security differs from
noninterference in a subtle but fundamental way, and approaches for
verifying noninterference cannot naturally extend to MPC security. Our
goal is to explore and establish connections between the security
model of MPC-- the \emph{real/ideal} aka \emph{simulator security}
model -- and trace-based hyperproperties, and to leverage these
connections to obtain automated enforcement mechanisms for MPC
protocol development. Currently, proof methods for MPC protocol
development are well-studied \cite{Lindell2017} but manual and thus
tedious and error-prone, and are also non-standardized and unfamiliar
to most PL theorists. Therefore our exploration will make both
theoretical and practical contributions by bridging a gap between
information flow and simulator security methodologies.

\subsection{The Analysis Challenge of MPC}
MPC protocols involve communication
between a group of distributed participants called a \emph{federation}
that collaboratively compute and publish the result of some known
\emph{ideal functionality} $\idealf$, while keeping each party's input
``secret'', without the use of a trusted third party. This last part
is critical. For example, if we take $\idealf$ to be the majority vote
function, a protocol for computing $\idealf$ is MPC-secure if, given
any set of input votes, it correctly computes and publishes the voting
result but reveals no other information to the public or to other
participants. However, by publishing the result, some information
about individual votes may be implicitly declassified.  For example,
in the case of a majority vote in a federation of size 3, if the
motion carries and party 1 has voted no, then party 1 knows exactly
the votes of parties 2 and 3. This cannot be avoided due to the nature
of $\idealf$.

@@

In Section \ref{section-hyperprop-passive} we formalize real/ideal
security, and we define and discuss simple examples of MPC protocols
in Section \ref{section-minicat-examples} and more complex ones in
Section \ref{section-metalang}. But an immediate and main point in
relation to security hyperproperties is that, due to the potential for
information release in MPC, simulator security is \emph{not} a strict
probabilistic noninterference or trace obliviousness property, as we
show in Section \ref{section-hyperprop-ni}- rather, the public output
allows and sets an upper bound on declassification.

@@

\paragraph{Language design.} In Section \ref{section-minicat} we
develop a new probabilistic programming language $\minifed$ for
defining synchronous distributed protocols over the binary field. The
syntax and semantics provides a succinct account of \emph{views} and
(synchronous) messaging between protocol \emph{clients}. In Section
\ref{section-metalang} we define a new metaprogramming language
$\metaprot$ that dynamically generates $\minifed$ protocols. It
includes control and data structures and is able to express logical
protocol components, and enjoys a type safety result that guarantees
that generated protocols are semantically well-defined.

\paragraph{Hyperproperty formulation.} In Section \ref{section-pmf} we
formulate \emph{program distributions}, a formalism for expressing
dependencies between secrets and views, useful for expressing
intensional characteristics of protocols and precisely quantifying
allowable information leakage in terms of protocol outputs. On this
basis we then formulate a novel hyperproperty of program execution
traces, called \emph{noninterference modulo output ($\NIMO$)} that we
demonstrate implies passive simulator security.

\paragraph{Fully and partially automated verification.} In Section
\ref{section-automation} we develop an automated technique, aka
certification, that verifies extensional properties of protocols such
as $\NIMO$, and also intensional properties of protocol
components. The latter include properties of protocol components that
provably guarantee security through composition in large
circuits. These compositional proofs are based on separation logic but
include novel results to support our semi-automated method for large
circuits where we (1) define certification for components, (2) certify
components, (3) prove $\NIMO$ based on the certifiedness of
components. We develop novel methods partially based on separation
logic \cite{barthe2019probabilistic} to enable (3).  In Section
\ref{section-composition} we illustrate our method to obtain
semi-automated verification methods for the GMW and Yao's Garbled
Circuit protocols, where certification can be re-used to validate new
gate implementations.
